---
title: "try_analysis_two"
author: "Xiaotong"
date: "14/08/2020"
output: html_document
---

```{r setup, include=FALSE}
options(width = 110)
library(tidyverse)
library(afex)
library(rtdists)
library(cowplot)
```


2. Method

3. results 
the questions i want to answer：
1."how good(rational) people are" (boxplot)


2. where are the mistakes (possibly) from? 


3."who is better/more rational" - numeracy, reasoning strategy, response time (run regression)


```{r}




### sanitise_ranks function
sanitise_ranks <- function(rank) {
  str_remove_all(rank, "_") %>%
    str_remove_all(" ")
}

### %!in% function
'%!in%' <- function(x, y) !('%in%'(x,y))


### experimental data
dat <- read.csv("ranking_data2.csv")

names(dat)[names(dat) == 'PROLIFIC_PID'] <- 'pid'

dl <- dat %>% 
  select(-ends_with("order")) %>% 
  pivot_longer(cols = starts_with("event_set"),
               names_to = "event", values_to = "response") %>% 
  filter(str_detect(response, "\\d")) %>% 
  mutate(set = str_extract(event, "\\d")) %>% 
  mutate(type = str_remove(event, "event_set\\d_")) %>%
  select(-event)

dl <- dl %>% 
  mutate(response = sanitise_ranks(response))


########
dl2 <- dat %>% 
  select(X, pid, ends_with("order")) %>% 
  pivot_longer(cols = starts_with("event_set"), 
               names_to = "event", values_to = "order") %>% 
  filter(!str_detect(order, "-")) %>%
  mutate(set = str_extract(event, "\\d")) %>%  
  select(-event) 

dl <- dl %>% 
  left_join(dl2)


########
 

dt_nest <- dl %>% 
  group_by(X, pid, condition, set, numeracy_score, response_time, strat_1, Strat_2, order ) %>%
  nest()



### coherent formalities
coherence_all <- read.csv("rank1_strict1_strict.csv", check.names=FALSE) 
coherence_all <- coherence_all[-1]
colnames(coherence_all) <- c("marginal", "conunction", "disjunction")


coherence_all <- coherence_all %>% 
  mutate(marginal = sanitise_ranks(marginal),
         conjunction = sanitise_ranks(conunction),
         disjunction = sanitise_ranks(disjunction))



### write the function to calculate the coherence of the beliefs about three event sets 
### （marginal event set, conjunction event set, and disjunction event set)


calc_logi_three <- function(data){
  ### browser()
  
  ### check if there are logical errors in rankings for marginal 
  if(data[data$type == "marginal", ]$response %!in%  coherence_all$marginal){
    return(1)
  } else {
    return(0)
  }
}


### try to run the function 
dt_nest <- dt_nest %>% 
  mutate(logical_err = map_dbl(data, calc_logi_three))


calc_coherence_all <- function(data){
  ### browser()
  
  ### check if there are logical errors in rankings for marginal 
  if(data[data$type == "marginal", ]$response %!in%  coherence_all$marginal){
    return(0) 
  }  ### if there are logical errors, return 0
  
  ### get possible rankings for conjunction given the rank in margianl 
  possible_ranks_conjunction <- coherence_all %>% 
    filter(marginal == data[data$type == "marginal", ]$response)
  
  #### check if response for conjunction matches 
  if(data[data$type == "conjunction", ]$response %!in% possible_ranks_conjunction$conjunction){
    return(0)
  }
  
  ### get possible rankings for disjunction given the ranks in marginal and conjunction 
  possible_ranks_disjunction <- coherence_all %>% 
    filter(marginal == data[data$type == "marginal", ]$response & 
             conjunction == data[data$type == "conjunction", ]$response)
  
  ### check if response for disjucntion matches 
  if(data[data$type == "disjunction", ]$response %in% possible_ranks_disjunction$disjunction){
    return(1)
  } else {
    return(0)
  }
  
}

dt_nest <- dt_nest %>% 
  mutate(coherence_all = map_dbl(data, calc_coherence_all))

calc_coh_marg_conj <- function(data){
  ### browser()
  
  ### check if there are logical errors in rankings for marginal 
  if(data[data$type == "marginal", ]$response %!in%  coherence_all$marginal){
    return(0) 
  }  ### if there are logical errors, return 0
  
  ### get possible rankings for conjunction given the ranks in margianl 
  possible_ranks_conjunction <- coherence_all %>% 
    filter(marginal == data[data$type == "marginal", ]$response)
  
  #### check if response for conjunction matches 
  if(data[data$type == "conjunction", ]$response %in% possible_ranks_conjunction$conjunction){
    return(1)
  } else{
    return(0)
  }
  
}

dt_nest <- dt_nest %>% 
  mutate(coh_marg_conj = map_dbl(data, calc_coh_marg_conj))

calc_coh_marg_disj <- function(data){
  ### check if there are logical errors in rankings for marginal 
  if(data[data$type == "marginal", ]$response %!in%  coherence_all$marginal){
    return(0) 
  }  ### if there are logical errors, return 0
  
  ### get possible rankings for disjunction given the ranks in margianl 
  possible_ranks_disjunction <- coherence_all %>% 
    filter(marginal == data[data$type == "marginal", ]$response)
  
  #### check if response for disjunction matches 
  if(data[data$type == "disjunction", ]$response %in% possible_ranks_disjunction$disjunction){
    return(1)
  } else{
    return(0)
  }
}


dt_nest <- dt_nest %>% 
  mutate(coh_marg_disj = map_dbl(data, calc_coh_marg_disj))

calc_coh_conj_disj <- function(data){
  
  ### get possible rankings for disjunction given the ranks in conjunction 
  possible_ranks_disjunction <- coherence_all %>% 
    filter(conjunction == data[data$type == "conjunction", ]$response)
  
  ### check if response for disjunction matches 
  if(data[data$type == "disjunction", ]$response %in% possible_ranks_disjunction$disjunction){
    return(1)
  }else{
    return(0)
  }
}

dt_nest <- dt_nest %>% 
  mutate(coh_conj_disj = map_dbl(data, calc_coh_conj_disj))


dt_nest <- dt_nest %>% 
  mutate(degree_of_coh = (coh_marg_conj + coh_marg_disj + coh_conj_disj) / 3) 

dt_nest[,12:16] %>% as.tibble() %>%
  print(n=40)

dt2 <- dt_nest %>% 
  select(-data)
```

Quesstion 1: how good people are ? 

- look at the distribution of the rate at which individual participants make logical errors under the full ranking condition
```{r}

dt2_full <- dt2 %>% filter(condition == "full") 


dt2_full <- dt2_full %>% group_by(pid) %>% 
  summarise( ave_logi = mean(logical_err),
             no_of_set = length(set))

ggplot(dt2_full, aes(x=ave_logi)) +
  geom_histogram(aes(y=stat(count/sum(count)/width)), binwidth = 0.1) + facet_wrap(~no_of_set) + 
  ylab("Density") + xlab("Rate of logical errors") + ggtitle("the distribution of the rate at which individual participants make logical errors under full ranking condition") +
  theme(plot.title = element_text(hjust = 0.5)) + theme_set(theme_bw(base_size = 9)) 



```

- look at the distribution of individual's degree of coherence under two conditions? 
```{r}


dt2_coherence <- dt2 %>% group_by(pid) %>% 
mutate(ave_coh = mean(degree_of_coh),
            no_of_set = length(set))


( boxplot1 <- ggplot(dt2_coherence, aes(x=condition, y=ave_coh)) +
  geom_boxplot() + 
  scale_y_continuous(name = "degree of coherence") +
        ggtitle("Boxplot of individual's degree of coherence (including responses with logical errors) ")  +
        theme_bw() +
        theme(plot.title = element_text(size = 10)))
  

dt2_coherence2 <- dt2_coherence %>% 
  filter(logical_err != "1")

( boxplot2 <- ggplot(dt2_coherence2, aes(x=condition, y=ave_coh)) +
  geom_boxplot() + 
  scale_y_continuous(name = "degree of coherence") +
        ggtitle("Boxplot of individual's degree of coherence (excluding responses with logical errors) ") +
        theme_bw() +
        theme(plot.title = element_text(size = 10)))

plot_grid(boxplot1, boxplot2) 



```


Question 2: where are the mistakes (possibly) from? 

method 1: "Does the manipulation of forcing participants to do pairwise comparisons for margianl events increase the coherence between conjunction and disjunction events?"
when using this method, i did not exclude any responses with logical errors

method 2: "Does the manipulation of forcing participants to do pairwise comparisons for marginal events increase the overall degree of coherence?" 
When using this method, i would run the ANOVA by including responses with logical errors and run the ANOVA by excluding responses with logical errors respectively.


```{r}
###--------- first step: add a column recording type of the event (weather event versus future related social events)

dt2$event_topic <- ifelse (dt2$set == "1" | dt2$set ==  "2" | dt2$set == "3" | dt2$set == "4", "weather", 
                          ifelse(dt2$set == "5" | dt2$set ==  "6" | dt2$set == "7" | dt2$set == "8", "social",
                                 "NA"))
                                                     
glimpse(dt2)


###-------DV: coherence between conjunction and disjunction events
a1 <- aov_ez("pid", "coh_conj_disj", dt2, between = "condition", within = "event_topic")
a1

afex_plot(a1, "condition", "event_topic", error = "within")
afex_plot(a1, "event_topic","condition",  error = "within") ### the effect of event_topic is significant 
                                                           ### the effect of condition is not significant

##---------- DV: degree of coherence, including responses with logical errors
a2 <- aov_ez("pid", "degree_of_coh", dt2, between = "condition", within = "event_topic")
a2

afex_plot(a2, "condition", "event_topic", error = "within") 
afex_plot(a2, "event_topic", "condition",  error = "within") ### the effect of event_topic is significant 
                                                           ### the effect of condition is not significant


##---------- DV: degree of coherence, excluding responses with logical errors
dt2_exclude <- dt2 %>% 
  filter(logical_err != "1")

a3 <- aov_ez("pid", "degree_of_coh", dt2_exclude, between = "condition", within = "event_topic")
a3

afex_plot(a3, "condition", "event_topic", error = "within") 
afex_plot(a3, "event_topic", "condition",  error = "within") ### the effect of event_topic is signiicant;
                                                            ### the effect of condition is significant

### participants who do not make logical errors when ranking margianl events under full ranking condition, on average, 
### perform better than participants under pairwise comparison condition in terms of the overall degree of coherence between three event sets (i.e. margianl, conjunction, and disjunction event sets)
### Besides, no matter what condition participants are in, they perform better on future related social events. 


### does the participants who do not make logical errors under full ranking condition perform better than those who do make logical errors under full ranking condition? 
### run regression 


dt2_full <- dt2 %>% 
  filter(condition == "full") %>% 
  group_by(pid) %>%
  mutate(ave_log_error = mean(logical_err))



ggplot(dt2_full, aes(x = ave_log_error, y = degree_of_coh, color = event_topic)) + 
  geom_point() + 
  geom_smooth(  method=lm) 

m1 <- lm(degree_of_coh ~ ave_log_error + event_topic + ave_log_error:event_topic, dt2_full) 
summary(m1) ### the coefficient of ave_log_error is significant


```

Does the the combination of event types matter
(For example, are people more likely to make errors (aka. behave against the probability theory) when they are judging "or" events and "and" events. 
In other words, will the "1" responses for conjunction and disjunction events be less than the "1" responses for marginal and conjunction/disjunction events? )

```{r}

dt2_two <- dt2 %>% 
  pivot_longer(starts_with("coh_"), names_to = "combination_of_two", values_to = "coherence_two") 

a1_coh_two <- aov_ez("pid", "coherence_two",dt2_two, between = "condition", within = c("combination_of_two", "event_topic"))
a1_coh_two
afex_plot(a1_coh_two, "combination_of_two", "condition", "event_topic", data_geom = geom_count) ###  the interaction condition:combination_of_two:event_topic is significant

emmeans::emmeans(a1_coh_two, "condition", by = c("combination_of_two", "event_topic")) %>% pairs() %>% update(by = NULL) %>% summary(adjust = "holm")

dt2_exclu_two <- dt2_exclude %>% 
   pivot_longer(starts_with("coh_"), names_to = "combination_of_two", values_to = "coherence_two") 

a2_coh_two <- aov_ez("pid", "coherence_two",dt2_exclu_two, between = "condition", within = c("combination_of_two", "event_topic"))
a2_coh_two
afex_plot(a2_coh_two, "combination_of_two", "condition", "event_topic") ###  the interaction condition:combination_of_two:event_topic is not significant

```

Considering the order of the events 

### ？？？

```{r}
dord_all <- dt2 %>% 
  group_by(pid, condition, order) %>% 
  summarise(across(coherence_all:degree_of_coh, mean)) %>% 
  separate(order, c("pos1", "pos2", "pos3"))

dord_all %>% 
  filter(pos1 == "marginal") %>% 
  filter(pos2 == "conjunction" | pos3 == "conjunction") %>% 
  ggplot(aes(x = pos2, y = coh_marg_conj)) +
  geom_boxplot() +
  stat_summary()

dord_all %>% 
  filter(pos1 == "marginal") %>% 
  filter(pos2 == "disjunction" | pos3 == "disjunction") %>% 
  ggplot(aes(x = pos2, y = coh_marg_disj)) +
  geom_boxplot() +
  stat_summary()

```


```{r}
dt2$marginal_order <- ifelse (dt2$order == "marginal_disjunction_conjunction" | dt2$order ==  "marginal_conjunction_disjunction" , "first", 
                          ifelse(dt2$order == "disjunction_marginal_conjunction" | dt2$order ==  "conjunction_marginal_disjunction" | dt2$order == "disjunction_conjunction_marginal" | dt2$order ==  "conjunction_disjunction_marginal", "later",
                                        "NA"))

dt2 %>% group_by(pid, marginal_order) %>% count()

dord <- dt2 %>% 
  group_by(pid, condition, marginal_order) %>%
  summarise(across(coherence_all:degree_of_coh, mean))


aov_ez("pid", "coh_marg_conj", dord, 
       between = "condition", within = c("marginal_order"))

aov_ez("pid", "coh_marg_disj", dord, 
       between = "condition", within = c("marginal_order"))

dt2_exclude$marginal_order <- ifelse (dt2_exclude$order == "marginal_disjunction_conjunction" | dt2_exclude$order ==  "marginal_conjunction_disjunction" , "first", 
                          ifelse(dt2_exclude$order == "disjunction_marginal_conjunction" | dt2_exclude$order ==  "conjunction_marginal_disjunction" | dt2_exclude$order == "disjunction_conjunction_marginal" | dt2_exclude$order ==  "conjunction_disjunction_marginal", "latter",
                                        "NA"))

dt2_exclude %>% group_by(pid, marginal_order) %>% count()


dt3 <- dt2 %>% filter()


```

Question 3: who is more rational? 
```{r}

dagg <- dt2 %>% 
  group_by(pid, condition) %>% 
  summarise(across(c(numeracy_score, response_time, coherence_all:degree_of_coh), mean))
afex::set_sum_contrasts()

m1 <- lm(coherence_all ~ numeracy_score*condition, dagg)
summary(m1)


## including logical errors


r1 <- lm(degree_of_coh ~ numeracy_score*event_topic*numeracy_score:event_topic*response_time, dt2) 
summary(r1)

### excluding logical errors 
r2 <- lm(degree_of_coh ~ numeracy_score*event_topic*numeracy_score:event_topic*response_time, dt2_exclude) 
summary(r2)

### under full ranking condition 
r3 <- lm(degree_of_coh ~ numeracy_score*event_topic*ave_log_error*response_time, dt2_full) 
summary(r3)


```


