---
title: "try_analysis_one"
author: "Xiaotong"
date: "13/08/2020"
output: html_document
---

# Calculate Dependent Variables

## Dependent variable one
Investigate if the responses for marginal, conjunction and disjunction event sets (aka. rankings for all types of event sets under the same topic) are coherent with each other. 
Record the coherent response as 1, and record the incoherent response as 0.

## Dependent variable two, three and four
Investigate if the responses for any two type of event sets (i.e. rankings for conjunction and marginal, rankings for disjunction and marginal, and rankings for conjunction and disjunction) are coherent with each other. 
Record the coherent response as 1, and record the incoherent response as 0.

## Dependent variable five: 
Calculate the degree of coherence for each participant. 
Definition of degree of coherence: (coherence between conjunction and margianl (0 or 1) + coherence between disjunction andm margianl (0 or 1) + coherence between conjunction and disjunction(0 or 1)) / 3 

## Depdendent variable six - logical error: 
Investigate if the participant will make logical errors when ranking marginal events under full ranking condition. 
Record the DV - Logical Error as 1 if the participant made the error, and record the DV - Logical Error as 0 if the participant did not make the error. 

# Question one
Is it necessary to exclude the responses with logical errors when analysing Dependent variable two, three, four and five, especially for coherence between conjunction and disjunction, and the degree of coherence?
For now, I exclude response with logical errors when analysing all DV other than DV six (logical error).

# Question two
Given some of the dependent variables are binary variables, will it be proper to do logistic regression? 
For now, I don't consider logistic regression.


```{r}
library(tidyverse)
library(afex)

### sanitise_ranks function
sanitise_ranks <- function(rank) {
  str_remove_all(rank, "_") %>%
    str_remove_all(" ")
}

### %!in% function
'%!in%' <- function(x, y) !('%in%'(x,y))


### experimental data
dat <- read.csv("ranking_data2.csv")

names(dat)[names(dat) == 'PROLIFIC_PID'] <- 'pid'

dl <- dat %>% 
  select(-ends_with("order")) %>% 
  pivot_longer(cols = starts_with("event_set"),
               names_to = "event", values_to = "response") %>% 
  filter(str_detect(response, "\\d")) %>% 
  mutate(set = str_extract(event, "\\d")) %>% 
  mutate(type = str_remove(event, "event_set\\d_")) %>%
  select(-event)

dl <- dl %>% 
  mutate(response = sanitise_ranks(response))


########
dl2 <- dat %>% 
  select(X, pid, ends_with("order")) %>% 
  pivot_longer(cols = starts_with("event_set"), 
               names_to = "event", values_to = "order") %>% 
  filter(!str_detect(order, "-")) %>%
  mutate(set = str_extract(event, "\\d")) %>%  
  select(-event) 

dl <- dl %>% 
  left_join(dl2)


########
 

dt_nest <- dl %>% 
  group_by(X, pid, condition, set, numeracy_score, response_time, strat_1, Strat_2, order ) %>%
  nest()



### coherent formalities
coherence_all <- read.csv("rank1_strict1_strict.csv", check.names=FALSE) 
coherence_all <- coherence_all[-1]


coherence_all <- coherence_all %>% 
  mutate(marginal = sanitise_ranks(`P(A)_P(¬A)_P(B)_P(¬B)`),
         conjunction = sanitise_ranks(`P(A^B)_P(A^¬B)_P(¬A^B)_P(¬B^¬A)`),
         disjunction = sanitise_ranks(`P(A∨B)_P(A∨¬B)_P(¬A∨B)_P(¬B∨¬A)`))



### write the function to calculate the coherence of the beliefs about three event sets 
### （marginal event set, conjunction event set, and disjunction event set)


calc_logi_three <- function(data){
  ### browser()
  
  ### check if there are logical errors in rankings for marginal 
  if(data[data$type == "marginal", ]$response %!in%  coherence_all$marginal){
    return(1)
  } else {
    return(0)
  }
}


### try to run the function 
dt_nest <- dt_nest %>% 
  mutate(logical_err = map_dbl(data, calc_logi_three))


calc_coherence_all <- function(data){
  ### browser()
  
  ### check if there are logical errors in rankings for marginal 
  if(data[data$type == "marginal", ]$response %!in%  coherence_all$marginal){
    return(0) 
  }  ### if there are logical errors, return 0
  
  ### get possible rankings for conjunction given the rank in margianl 
  possible_ranks_conjunction <- coherence_all %>% 
    filter(marginal == data[data$type == "marginal", ]$response)
  
  #### check if response for conjunction matches 
  if(data[data$type == "conjunction", ]$response %!in% possible_ranks_conjunction$conjunction){
    return(0)
  }
  
  ### get possible rankings for disjunction given the ranks in marginal and conjunction 
  possible_ranks_disjunction <- coherence_all %>% 
    filter(marginal == data[data$type == "marginal", ]$response & 
             conjunction == data[data$type == "conjunction", ]$response)
  
  ### check if response for disjucntion matches 
  if(data[data$type == "disjunction", ]$response %in% possible_ranks_disjunction$disjunction){
    return(1)
  } else {
    return(0)
  }
  
}

dt_nest <- dt_nest %>% 
  mutate(coherence_all = map_dbl(data, calc_coherence_all))

calc_coh_marg_conj <- function(data){
  ### browser()
  
  ### check if there are logical errors in rankings for marginal 
  if(data[data$type == "marginal", ]$response %!in%  coherence_all$marginal){
    return(0) 
  }  ### if there are logical errors, return 0
  
  ### get possible rankings for conjunction given the ranks in margianl 
  possible_ranks_conjunction <- coherence_all %>% 
    filter(marginal == data[data$type == "marginal", ]$response)
  
  #### check if response for conjunction matches 
  if(data[data$type == "conjunction", ]$response %in% possible_ranks_conjunction$conjunction){
    return(1)
  } else{
    return(0)
  }
  
}

dt_nest <- dt_nest %>% 
  mutate(coh_marg_conj = map_dbl(data, calc_coh_marg_conj))

calc_coh_marg_disj <- function(data){
  ### check if there are logical errors in rankings for marginal 
  if(data[data$type == "marginal", ]$response %!in%  coherence_all$marginal){
    return(0) 
  }  ### if there are logical errors, return 0
  
  ### get possible rankings for disjunction given the ranks in margianl 
  possible_ranks_disjunction <- coherence_all %>% 
    filter(marginal == data[data$type == "marginal", ]$response)
  
  #### check if response for disjunction matches 
  if(data[data$type == "disjunction", ]$response %in% possible_ranks_disjunction$disjunction){
    return(1)
  } else{
    return(0)
  }
}


dt_nest <- dt_nest %>% 
  mutate(coh_marg_disj = map_dbl(data, calc_coh_marg_disj))

calc_coh_conj_disj <- function(data){
  
  ### get possible rankings for disjunction given the ranks in conjunction 
  possible_ranks_disjunction <- coherence_all %>% 
    filter(conjunction == data[data$type == "conjunction", ]$response)
  
  ### check if response for disjunction matches 
  if(data[data$type == "disjunction", ]$response %in% possible_ranks_disjunction$disjunction){
    return(1)
  }else{
    return(0)
  }
}

dt_nest <- dt_nest %>% 
  mutate(coh_conj_disj = map_dbl(data, calc_coh_conj_disj))


dt_nest <- dt_nest %>% 
  mutate(degree_of_coh = (coh_marg_conj + coh_marg_disj + coh_conj_disj) / 3) 

dt_nest[,12:16] %>% as.tibble() %>%
  print(n=40)



```

```{r}
dt_nest %>% 
  group_by(degree_of_coh) %>% 
  count()

dt_nest %>% 
  filter(condition == "full") %>%
  group_by(logical_err) %>%
  count()

```
### 1: analyse dependent variable two, three and four （i.e. responses for any two type of event sets (i.e. rankings for conjunction and marginal, rankings for disjunction and marginal, and rankings for conjunction and disjunction) 
### and control some IVs 


-  main question:
Does the the combination of event types matter?
(For example, are people more likely to make errors (aka. behave against the probability theory) when they are judging "or" events. 
In other words, will the "1" responses for marginal and "or" events be less than the "1" responses for marginal and "and" events? )

### whith-subject: the combination of event types

### within subject: order
Does the order matter? -within subjet
- the order within two event set. For example, when ranking marginal and conjunction event sets, wether the participant see margianl or conjunction first will influence the coherence between marginal and conjunction. 
- the order of presentation of two event set. For example, whether the participant ranking the marginal and conjunction event sets before，after or apart (i.e. separated by the disjunction event set) will influence the coherence between margianl and conjunction.


### within subject: topic or set no.
Does the topic matter (whether and future related social events)?  

### between subject: condition
Does the condition matter (i.e. full ranking and pairwise comparison condition) 

### question: how to analyse IV "order" ? 
### question: how to analyse IV "topic". Topic is not a "pure" within or between subject. For example some particiapnts saw event set 1, 3, 4, 7; while some participants saw event set 2, 3, 4, 8. So there are only some overlaps (or sometimes no overlaps) on topics they were assigned. The same question applies to order
- do regression analysis? if so, what regression ? 


```{r}
## step 1: exclude responses with logical error 

dt_nest2 <- dt_nest %>% 
  filter(logical_err != "1")


dt_nest2 <- dt_nest %>% 
  pivot_longer(starts_with("coh_"), names_to = "combination_of_two", values_to = "coherence_two") %>% 
  select(-data)

```

### For now, I set up a model with condition as between subject variable and combination_of_two_as within subject variable.
```{r}
### repeated measures ANOVA
### set up main model
### DV:coherence_two  (binary varible: 1 or 0)
### within subject: combination_of_two, topic(or set), order
### between subject: condition



a1_coh_two <- aov_ez("pid", "coherence_two", dt_nest2, between = "condition", within = c("combination_of_two"))

a1_coh_two

afex_plot(a1_coh_two, "condition", "combination_of_two", error = "within")

afex_plot(a1_coh_two, "combination_of_two", "condition", error = "within" )
### group the dataset according to condition

```
###2.DV: degree of coherence

###: within subject
Does the order matter? - （～within?)  subjet
Does the topic matter (whether and future related social events)?  - (~within?) subject
### questions: same questions as before when it comes to order and topic. 

### between subject
Does the condition matter (i.e. full ranking and pairwise comparison condition) 


```{r}

```

###3.DV: degree of coherence (three level, 0.3333, 0.6667 and 1)

### IV: (all between subjects?)
Dpes the numeracy skills matter

Does the reasoning strategy matter 

Does the response time matter?

### Question: do i need to consider the education?

### question: how to convert the response time to factor with levels?


```{r}
### first step: factorise the variable: numeracy skills


dt_nest$numeracy_skill <- ifelse (dt_nest$numeracy_score %in% seq(0, 2, 1), "low", 
                                   ifelse(dt_nest$numeracy_score %in% seq(3, 5, 1), "medium",
                                          ifelse(dt_nest$numeracy_score %in% seq(6, 7, 1), "high", 
                                                 "NA")))

dt_nest <- dt_nest %>% select(-data) 
dt_nest <- dt_nest %>% select(-1)
dt_nest <- dt_nest %>% unique()

dt_nest$numeracy_skill <- dt_nest$numeracy_skill %>%  as.factor()

### second step: turn the strategy questions into factors

dt_nest$strat_1 <- dt_nest$strat_1 %>% as.factor()
dt_nest$Strat_2 <- dt_nest$Strat_2 %>% as.factor()
glimpse(dt_nest)


### three-way ANOVA 
### set up the main model
### DV： degree_of_coh
### between subject IV: numeracy skill, strat_1, Strat_2, response time
a1_degree_coh <- aov_ez("pid", "degree_of_coh", dt_nest, between = c("numeracy_skill", "condition" ))

a1_degree_coh



```

###4.DV: logical error 

### IV: 
order
topic
numeracy
reasoning strategy
duration



